Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-02-05 arxiv/from-weak-separability-to-strong-separability.tex
added arxiv folder

We use a positive definite kernel function $k(\cdot, \cdot)$, namely
\emph{rational kernel}~\citep{Shalev-Shwartz-Shamir-Sridharan-2011}, whose
corresponding feature map $\phi(\cdot)$ transforms any sequence of \emph{weakly}
linearly separable examples to \emph{strongly} linearly separable sequence of
examples. Specifically, $\phi$ has the property that if a set of labeled
examples in $\B(\zero,1)$ is weakly linearly separable with a margin $\gamma$, then
after applying $\phi$ the examples become strongly linearly separable with a
margin $\gamma'$ and their squared norms are bounded by $2$. The parameter $\gamma'$ is
a function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(\zero,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
%\begingroup
%\allowdisplaybreaks
%\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
%\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
%& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
%\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(4K-5) 2^{K-1}} \\
%& \ \cdot \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
%\end{align*}
%\endgroup
\begingroup
\allowdisplaybreaks
\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 = & \frac{
  \left[
    376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil
  \right]^{
    \frac{-\lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{{2}/{\gamma}} \right \rceil}{2}
  }
}{2\sqrt{K}},
\\
\gamma_2 = & \frac{
   \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)}
}{4\sqrt{K}(4K-5) 2^{K-1}}
\\
&  \cdot 2^{s(s+1)r(K-1)}
\; ,
\end{align*}
\endgroup
where $r = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1$ and $s = \left \lceil \log_2(2/\gamma) \right \rceil$.
Then, the sequence of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, for all $t$ in $\cbr{1,\ldots,T}$,
$k(x_t,x_t) \leq 2$.
\end{theorem}

@@

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive integer and let $\gamma$ be a positive real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \B(\zero,1) \times \{1,2,\dots,K\}$
is a sequence of weakly separable labeled examples with margin $\gamma > 0$,
then the expected number of mistakes made by Algorithm~\ref{algorithm:kernelized}
with kernel $k(x,x')$ defined by \eqref{equation:rational-kernel}
is at most $2^{\widetilde{O}(\min(K \log^2 \frac{1}{\gamma},
\sqrt{\frac{1}{\gamma}} \log K))}$.
\end{corollary}

@@

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, \qquad p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with margin $\gamma'$.

@@

\item $\deg(p) \le (2s+1) rm$,
\item $\norm{p} \le (4m-1) 2^m \cdot \left(2^s rm (4s+2)^2 \right)^{(s+1/2)rm}$.
\end{enumerate}
\end{theorem}

@@

Using the lemma and the polynomial approximation theorems, we can prove that the
mapping $\phi$ maps any set of weakly linearly separable examples to a strongly
linearly separable set of examples. Due to space constraints, we defer the full
proof to Appendix~\ref{section:proof-of-theorem-margin-transformation}.

>>>>

We use a positive definite kernel function $k(\cdot, \cdot)$, namely
\emph{rational kernel}~\citep{Shalev-Shwartz-Shamir-Sridharan-2011}, whose
corresponding feature map $\phi(\cdot)$ transforms any sequence of \emph{weakly}
linearly separable examples to \emph{strongly} linearly separable sequence of
examples.
Specifically, $\phi$ has the property that if a set of labeled
examples in $\B(\zero,1)$ is weakly linearly separable with a margin $\gamma$, then
after applying $\phi$ the examples become strongly linearly separable with a
margin $\gamma'$ and their squared norms are bounded by $2$.
\footnote{Other kernels, such as the polynomial kernel
$k(x,x') = (1+\ip{x}{x'})^d$, or the multinomial kernel~\citep{Goel-Klivans-2017} $k(x,x') = \sum_{i=0}^d(\ip{x}{x'})^i$,
will have similar properties for large enough $d$.}
The parameter $\gamma'$ is
a function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(\zero,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
%\begingroup
%\allowdisplaybreaks
%\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
%\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
%& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
%\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(4K-5) 2^{K-1}} \\
%& \ \cdot \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
%\end{align*}
%\endgroup
\begingroup
\allowdisplaybreaks
\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 = & \frac{
  \left[
    376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil
  \right]^{
    \frac{-\lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{{2}/{\gamma}} \right \rceil}{2}
  }
}{2\sqrt{K}},
\\
\gamma_2 = & \frac{
   \left(2^{s+1} r(K-1) (4s+2) \right)^{-(s+1/2)r(K-1)}
}{4\sqrt{K}(4K-5) 2^{K-1}}
\; ,
\end{align*}
\endgroup

%\\
%&  \cdot 2^{s(s+1)r(K-1)}

where $r = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1$ and $s = \left \lceil \log_2(2/\gamma) \right \rceil$.
Then, the sequence of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, for all $t$ in $\cbr{1,\ldots,T}$,
$k(x_t,x_t) \leq 2$.
\end{theorem}

@@

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive integer and let $\gamma$ be a positive real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \B(\zero,1) \times \{1,2,\dots,K\}$
is a sequence of weakly separable labeled examples with margin $\gamma > 0$,
then the expected number of mistakes made by Algorithm~\ref{algorithm:kernelized}
with kernel $k(x,x')$ defined by \eqref{equation:rational-kernel}
is at most $\min (2^{\widetilde{O}(K \log^2
(1/\gamma))}, 2^{\widetilde{O}(\sqrt{1/\gamma} \log K)})$.
\end{corollary}

@@

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, \qquad p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with a positive margin.

@@

\item $\deg(p) \le (2s+1) rm$,
\item $\norm{p} \le (4m-1) 2^m \cdot \left(2^s rm (4s+2) \right)^{(s+1/2)rm}$.
\end{enumerate}
\end{theorem}

@@

Using the lemma and the polynomial approximation theorems, we can prove that the
mapping $\phi$ maps any set of weakly linearly separable examples to a strongly
linearly separable set of examples. Due to space constraints, we defer the full
proof of Theorem~\ref{theorem:margin-transformation} to
Appendix~\ref{section:proof-of-theorem-margin-transformation}.

### 2019-02-05 arxiv/notions-of-linear-separability.tex
added arxiv folder

We define two notions of linear separability for multiclass classification. The
first notion is the standard notion of linear separability used in the proof of
the mistake bound for the \textsc{Multiclass Perceptron} algorithm. The second
notion is stronger, i.e. more restrictive. %However, it is more suitable for the
%bandit setting, since it allows for a simple and efficient algorithm; see
%Section~\ref{section:algorithm-for-strongly-linearly-separable-data}.

>>>>

We define two notions of linear separability for multiclass classification. The
first notion is the standard notion of linear separability used in the proof of
the mistake bound for the \textsc{Multiclass Perceptron} algorithm\citep[see e.g.][]{Crammer-Singer-2003}. The second
notion is stronger, i.e. more restrictive. %However, it is more suitable for the
%bandit setting, since it allows for a simple and efficient algorithm; see
%Section~\ref{section:algorithm-for-strongly-linearly-separable-data}.

### 2019-02-05 arxiv/title-and-abstract.tex
added arxiv folder

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case}

\twocolumn[
\icmltitle{Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case}

\begin{icmlauthorlist}
\icmlauthor{Alina Beygelzimer}{yahoo}
\icmlauthor{D\'avid P\'al}{yahoo}
\icmlauthor{Bal\'azs Sz\"or\'enyi}{yahoo}
\icmlauthor{Devanathan Thiruvenkatachari}{nyu}
\icmlauthor{Chen-Yu Wei}{usc}
\icmlauthor{Chicheng Zhang}{microsoft}

\end{icmlauthorlist}

\icmlaffiliation{yahoo}{Yahoo Research, New York, NY, USA}
\icmlaffiliation{nyu}{New York University, New York, MA, USA}
\icmlaffiliation{usc}{University of Southern California, Los Angeles, CA, USA}
\icmlaffiliation{microsoft}{Microsoft Research, New York, NY, USA}

\icmlcorrespondingauthor{D\'avid P\'al}{davidko.pal@gmail.com}

\icmlkeywords{multi-armed bandits, contextual bandits, online classification, linear separability}

\vskip 0.3in
]

\printAffiliationsAndNotice{}

>>>>

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
%\icmltitlerunning{Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case}

%\twocolumn[
%\icmltitle{Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case}

\title{Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case}

%\begin{icmlauthorlist}
\author[1]{Alina Beygelzimer}
\author[1]{D\'avid P\'al}
\author[1]{Bal\'azs Sz\"or\'enyi}
\author[2]{Devanathan Thiruvenkatachari}
\author[3]{Chen-Yu Wei}
\author[4]{Chicheng Zhang}

%\end{icmlauthorlist}

\affil[1]{Yahoo Research, New York, NY, USA}
\affil[2]{New York University, New York, MA, USA}
\affil[3]{University of Southern California, Los Angeles, CA, USA}
\affil[4]{Microsoft Research, New York, NY, USA}

\maketitle
%\icmlcorrespondingauthor{D\'avid P\'al}{davidko.pal@gmail.com}

%\icmlkeywords{multi-armed bandits, contextual bandits, online classification, linear separability}

%\vskip 0.3in
%]

%\printAffiliationsAndNotice{}
